National Longitudinal Evidence for Growth in Subjective Well-being from Spiritual Beliefs

### 2021-02-12 Supplement_And_Analysis.Rmd
fixed bib

# Bayesian counfounder model

>>>>

## Priors in the Bayesian Model

### Priors for the Personal Wellbeing model

Model priors for Personal Wellbeing including group-level ID intercept and beliefs slopes are as follows:

$$
\begin{aligned}
\operatorname{Personal Well Being}_{ij} \sim \text{Normal}(\mu_{ij},\sigma^2) \nonumber\\
\mu_{ij}\sim \boldsymbol{\alpha +\beta} \nonumber\\
\boldsymbol{\alpha} \sim \alpha_{0} +\alpha_j \nonumber\\
\alpha_{0} \sim \text{StudentT}(3,7.2,2.5)\nonumber\\
\boldsymbol{\beta}\sim \boldsymbol{b} + \beta_j \nonumber\\
\boldsymbol{b}\sim\text{Normal}(0,2.5) \nonumber\\
\begin{bmatrix}
\alpha_j \\
\beta_j
\end{bmatrix}
\sim 
\text{MVNormal}
\begin{pmatrix}
\begin{bmatrix}
0\\
0
\end{bmatrix}
,\boldsymbol{S} \nonumber
\end{pmatrix}\\
\boldsymbol{S} = 
\begin{pmatrix}
0 & \sigma_{\beta_j} \nonumber
\end{pmatrix} \boldsymbol{R} \begin{pmatrix}
\sigma_{\alpha_j} & 0 \\
0 & \sigma_{\beta_j} \nonumber
\end{pmatrix}\\
\sigma_{\alpha_j} \sim 
\text{HalfCauchy}(0,2.5) \nonumber\\
\sigma_{\beta_j} \sim 
\text{HalfCauchy}(0,2.5) \nonumber\\
\boldsymbol{R}\sim \text{LKJcorr}(1) \nonumber\\
\sigma  \sim  \operatorname{HalfStudent T}^{+}(3, 0, 2.5) 
\end{aligned}
$$

### Priors for the Life Satisfaction model

Model priors for Life Satisfaction Model with including group-level ID intercept and belief slopes are as follows:

$$
\begin{aligned}
\operatorname{Life Satisfaction}_{ij} \sim \text{Normal}(\mu_{ij},\sigma^2) \nonumber\\
\mu_{ij}\sim \boldsymbol{\alpha +\beta} \nonumber\\
\boldsymbol{\alpha} \sim \alpha_{0} +\alpha_j \nonumber\\
\alpha_{0} \sim \text{StudentT}(3,5.5,2.5)\nonumber\\
\alpha_j \sim \text{StudentT}+(3,0,2.5)\nonumber\\
\boldsymbol{\beta}\sim \boldsymbol{b} + \beta_j \nonumber\\
\boldsymbol{b}\sim\text{Normal}(0,2.5) \nonumber\\
\begin{bmatrix}
\alpha_j \\
\beta_j
\end{bmatrix}
\sim 
\text{MVNormal}
\begin{pmatrix}
\begin{bmatrix}
0\\
0
\end{bmatrix}
,\boldsymbol{S} \nonumber
\end{pmatrix}\\
\boldsymbol{S} = 
\begin{pmatrix}
\sigma_{\alpha_j} & 0 \\
0 & \sigma_{\beta_j} \nonumber
\end{pmatrix} \boldsymbol{R} \begin{pmatrix}
\sigma_{\alpha_j} & 0 \\
0 & \sigma_{\beta_j} \nonumber
\end{pmatrix}\\
\sigma_{\alpha_j} \sim 
\text{HalfCauchy}(0,1) \nonumber\\
\sigma_{\beta_j} \sim 
\text{HalfCauchy}(0,1) \nonumber\\
\boldsymbol{R}\sim \text{LKJcorr}(1) \nonumber\\
\sigma  \sim  \operatorname{HalfStudent T}^{+}(3, 0, 2.5) 
\end{aligned}
$$
\newpage

### Deterministic model

The deterministic portion of both the Personal Wellbeing and the Life Satisfaction models is as follows: 
$$
\begin{aligned}
& \quad  \boldsymbol{b} = \\
& \quad \beta_{1}(\operatorname{Beliefs}_{\operatorname{SpiritExcludesGod}})\ + \\
&\quad \beta_{2}(\operatorname{Beliefs}_{\operatorname{GodAndSpirit}})\  + \\
& \quad  \beta_{3}(\operatorname{Beliefs}_{\operatorname{GodExcludesSpirit}})\ +\\
&\quad \beta_{4}(\operatorname{Age in Decades (C)})\  +\\ 
& \quad  \beta_{5}(\operatorname{Ethnicity}_{\operatorname{Maori}})\ +\\
&\quad \beta_{6}(\operatorname{Ethnicity}_{\operatorname{Pacific}})\ +  \\ 
& \quad  \beta_{7}(\operatorname{Ethnicity}_{\operatorname{Asian}})\ +\\
& \quad \beta_{8}(\operatorname{Male} )+ \\ 
& \quad \beta_{9}(\operatorname{Years} \times \operatorname{Beliefs}_{\operatorname{SpiritExcludesGod}})\ + \\ 
& \quad \beta_{10}(\operatorname{Years} \times \operatorname{Beliefs}_{\operatorname{GodAndSpirit}})\ + \\
&\quad \beta_{11}(\operatorname{Years} \times \operatorname{Beliefs}_{\operatorname{GodExcludesSpirit}}) \\ 
\end{aligned}
$$
\newpage

## Bayesian counfounder model
